---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# topicsplorrr

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/sdaume/topicsplorrr/workflows/R-CMD-check/badge.svg)](https://github.com/sdaume/topicsplorrr/actions)
[![Codecov test coverage](https://codecov.io/gh/sdaume/topicsplorrr/branch/master/graph/badge.svg)](https://codecov.io/gh/sdaume/topicsplorrr?branch=master)
<!-- badges: end -->

The package provides functions to analyse and visualize the results of an [STM](https://github.com/bstewart/stm) topic model, including a Shiny application for topic model exploration.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("sdaume/topicsplorrr")
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r eval=FALSE}
library(topicsplorrr)
## basic example code
```



## License, credits and acknowledgements
The package is shared under an [MIT License](LICENSE.md). 

This package has been developed to support research at the [Stockholm Resilience Centre](https://www.stockholmresilience.org); this research has benefited from funding by the [Swedish Research Council for Sustainable Development (Formas)](https://formas.se/).


## Disclaimer
This package has been developed as a reusable tool for the author(s) own research and comes with no guarantee for the correctness of the supported analyses. 

