---
title: "Simple analysis of term frequencies"
output: 
  tufte::tufte_html:
    tufte_variant: envisioned
    toc: true
    css: vignette.css
    margin_references: false
vignette: >
  %\VignetteIndexEntry{Simple analysis of term frequencies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  cache = FALSE,
  comment = "#>"
)
```

```{r setup}
library(topicsplorrr)
```


## "Naive" topic analysis

The package primarily supports provides functions to analyze and visualize the results of an `stm` topical model, but includes also functions to explore simple term trends in a text corpus. This can be thought of as a first "naive" topical analysis and is a useful initial exploration of a text collection.

All examples in the following sections are using a snapshot of the [*'SARS-CoV-2 COVID-19'* preprint collection](https://connect.medrxiv.org/relate/content/181) on [medrXiv](https://www.medrxiv.org/) and [biorXiv](https://www.biorxiv.org/); this data is available via the [`covid19topics` package](https://github.com/sdaume/covid19topics) which has been setup for a topical analysis.

For the following analyses common English stop words were typically removed from the source texts, as are numbers and words were ["stemmed"](https://en.wikipedia.org/wiki/Stemming)^[I.e. all words are reduced to their word stem and words like *"model", "models", "modelling", "modeling"* are transformed to the word stem *"model"*].  

The following table utilizes the package's `term_counts()` function to list the most frequent terms in the analyzed preprint titles.


```{r echo=TRUE}
library(topicsplorrr)
library(dplyr)

# extract unigrams
processed_terms <- unigrams_by_date(textData = covid19topics::covid19preprints, 
                                    textColumn = "title", 
                                    dateColumn = "pub_date")

# and compute term shares
top_title_terms <- term_counts(processed_terms) %>%
  slice(1:10) %>%
  mutate(term_share = term_share*100)

top_title_terms %>% 
  kableExtra::kbl(format = "html", 
                  caption = "Most frequent terms in Covid19 preprint titles",
                  col.names = c("Term", "N", "%"),
                  digits = c(0,0,2)) 
```

