---
title: "Simple analysis of STM topic models"
output: 
  html_document:
    toc: false
bibliography: references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{Simple analysis of STM topic models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  cache = FALSE,
  comment = "#>"
)
```

```{r setup}
library(topicsplorrr)
library(stm)
library(dplyr)
library(quanteda)
library(DT)
library(extrafont)

# use quanteda's inaugural presidential speeches as sampel data
sample_docs <- quanteda::convert(quanteda::data_corpus_inaugural,
                                        to = "data.frame") %>%
  mutate(Year = lubridate::as_date(paste(Year, "-01-20", sep = "")))

# setup variables used in the vignette examples
topic_labels <- topics_terms_map(sample_topics)
topic_doc_shares <- sample_topic_doc_probs
```


## Fit an STM topic model to a document collection
The `topicsplorrr()` package currently supports analyses of *structural topic models* created with the the versatile `stm` package [@RobertsStewart_et_2014_AJPS_58]. `stm` builds on the tradition of popular topic modeling approaches like *Latent Dirichlet Allocation* [@BleiNgEt2003_JMLR_3], but allows to incorporate other document metadata into the topic model [@Roberts_et_al_2019_JSTATS_91_2].

In order to explore an `stm` topic model with the help of the `topicsplorrr()` package a topic model needs to be fit and information on topic-document likelihoods and word-topic likelihoods needs to be extracted. The code below prepares a document collection, fits an `stm` topic model to this.  

Consult the respective package documentation for more information on the `terms_dfm()` and `stm::stm()` methods. For later analyses it is also assumed that the original document data contains a date for each document.


```{r, eval=FALSE, echo=TRUE}
library(topicsplorrr)
library(stm)
library(dplyr)
library(lubridate)

# a dataframe with documents 
# use quanteda's inaugural presidential speeches as sampel data
sample_docs <- quanteda::convert(quanteda::data_corpus_inaugural,
                                        to = "data.frame") %>%
  mutate(Year = lubridate::as_date(paste(Year, "-01-20", sep = "")))

# a document feature matrix
sample_dfm <- terms_dfm(textData = sample_docs, 
                        textColumn = "text", documentIdColumn = "doc_id", 
                        removeStopwords = TRUE, removeNumbers = TRUE, 
                        wordStemming = FALSE)

# fitting an STM topic model
sample_topics <- stm::stm(sample_dfm, K = 25)
```


## Topic likelihoods and labels
The `stm()` package provides several functions to explore the topic model. The `plot` function provides an overview of the topic model with expected topic proportions for the whole document collection and with the five most probable words for each topic.

```{r, fig.height=8, fig.width=10, dpi=300, eval=TRUE, echo=TRUE}
plot(sample_topics, type = "summary", n = 5)
```

The `topicsplorrr` package provides some convenience methods to extract the same information for customized analyses and visualizations. With the help of the `topics_by_doc_date` function the topic likelihoods (or "shares") for each document can be obtained. Below an example for the document set used so far.

```{r eval=FALSE, echo=TRUE}
topic_doc_shares <- topics_by_doc_date(sample_topics, 
                                       termsDfm = sample_dfm, 
                                       textData = sample_docs,
                                       documentIdColumn = "doc_id", 
                                       dateColumn = "Year")
```

Resulting in a dataframe with the following variables:

```{r eval=TRUE, echo=TRUE}
dplyr::glimpse(topic_doc_shares)
```

Useful topic labels can be generated with the help of most likely words associated with a topic. The following method is a simple wrapper around methods in the `stm()` package. As a  default it returns the seven terms that are both most frequent and exclusive for a topic identified by ID. 

```{r eval=FALSE, echo=TRUE}
topic_labels <- topics_terms_map(sample_topics)
```

This information can be combined to create alternative visualizations, for example an interactive `DT::datatable`.

```{r eval=FALSE, echo=TRUE}  
topic_means <- topics_summary(topic_doc_shares, topic_labels)
  
```

```{r eval=TRUE}  
topic_means <- topics_summary(topic_doc_shares, topic_labels)

DT::datatable(topic_means, 
          options = list(pageLength = 10, 
                         columnDefs = list(list(className = 'dt-left', 
                                                targets = "_all"))), 
          rownames = FALSE, 
          colnames = c("Topic ID", "Most representative words", "Mean topic probability")) %>%
  DT::formatRound(3, digits = 4) %>%
  DT::formatStyle(columns = c(1, 2, 3), fontSize = '90%') %>%
  DT::formatStyle(columns = c(1, 3), 'text-align' = 'right') %>%
  #DT::formatStyle(columns = c(1), width = '10%') %>%
  #DT::formatStyle(columns = c(2), width = '60%') %>%
  DT::formatStyle(3,
              background = DT::styleColorBar(range(topic_means$mean_gamma), '#1964A6'),
              backgroundSize = '100% 88%',
              backgroundRepeat = 'no-repeat',
              backgroundPosition = 'center')
```


## Temporal topic trends

One of the main objectives of `topicsplorrr` is to analyze temporal trends of topic models. The `topics_by_doc_date()` function extracts the topic likelihoods for each document-topic combination and merges this with the original document information, specifically the document identifiers and document publication date.

```{r, eval=FALSE, echo=TRUE}
topic_doc_shares <- topics_by_doc_date(sample_topics, 
                                       termsDfm = sample_dfm, 
                                       textData = sample_docs,
                                       documentIdColumn = "doc_id", 
                                       dateColumn = "Year")
```

These can be used for custom analysis or a direct visualization of topical trends combining `topics_terms_map()` and `plot_topic_frequencies()`. 


```{r eval=FALSE, echo=TRUE, fig.width=11, fig.height=11, dpi=300}
# create a topic label set
topic_labels <- topicsplorrr::topics_terms_map(sample_topics, labelType = "prob")

# plot frequencies for all topics
topic_doc_shares %>%
  plot_topic_frequencies(topN = 16, nCols = 4, selectTopicsBy = "most_frequent", 
                         timeBinUnit = "year", topicLabels = topic_labels, 
                         verboseLabels = FALSE)
```


## Topic signatures
TBD


## References
